\textbf{How to measure maximally precise radial velocities from a set of spectra:\\ The time-independent spectrum case}

### 2019-08-26 paper/eprv.tex
tweaks

\CRLB-saturating estimators are (or are equivalent to)
maximum-likelihood estimators.
When the likelihood is Gaussian, the log-likelihood
is a sum of a quadratic chi-squared-like term and a determinant term
\begin{equation}
\ln L = -\frac{1}{2}\,[y - f(v)]\T\cdot C^{-1}\cdot [y - f(v)] - \frac{1}{2}\,||2\pi\,C|| \quad ,
\label{eq:lf}
\end{equation}
where $y$ is a column vector of the data, $f(v)$ is the model spectrum,
which depends on the \RV\ $v$,
and $C$ is the covariance matrix describing the noise variance.
In the case of independent pixel noise the chi-squared term just becomes an
inverse-variance weighted sum of squared residuals (chi-squared).
The maximum-likelihood estimator for $v$ is the value of $v$ which maximizes
the log-likelihood.
Since only the chi-squared term depends on $v$, maximizing the likelihood
means minimizing chi-squared.

Most methods for radial-velocity estimation aren't explicitly maximizing
any likelihood.
Instead, they involve some kind of
cross-correlation of the data with a template.
This might be surprising, since (as we noted above) bound-saturating
estimators will be maximum-likelihood estimators.
However, under simple assumptions, cross-correlation optima are
also maximum-likelihood optima. The argument is as follows:
If we expand the quadratic form (chi-squared term) in the
likelihood in equation~(\ref{eq:lf}), it becomes
\begin{equation}
{}-\frac{1}{2}\,y\T\cdot C^{-1}\cdot y + y^T\cdot C^{-1}\cdot f(v) 
-\frac{1}{2}\,f(v)\T\cdot C^{-1}\cdot f(v) \quad.
\end{equation}
That is, the chi-squared term factors into a term that is the square
of the model, a term that is the square of the data, and a term that
is a cross-correlation.
As long as the amplitude of $f$ doesn't depend on velocity, and the
$C$ tensor is ``smooth'' in certain respects, the dependence of this
chi-squared term on velocity will be dominated by the
cross-correlation.
For this reason, we can use a variant of cross-correlation
optimization as a radial-velocity estimator.

>>>>

\CRLB-saturating estimators are (or are equivalent to)
maximum-likelihood estimators.
When the likelihood is Gaussian, the log-likelihood
is a sum of a quadratic chi-squared-like term and a log-determinant term
\begin{equation}
\ln L = -\frac{1}{2}\,[y - f(v)]\T\cdot C^{-1}\cdot [y - f(v)] - \frac{1}{2}\,\ln||2\pi\,C|| \quad ,
\label{eq:lf}
\end{equation}
where $y$ is a column vector of the data, $f(v)$ is the model spectrum,
which depends on the \RV\ $v$,
and $C$ is the covariance matrix describing the noise variance.
In the case of independent pixel noise the chi-squared term just becomes an
inverse-variance weighted sum of squared residuals (chi-squared).
The maximum-likelihood estimator for $v$ is the value of $v$ which maximizes
the log-likelihood.
Since only the chi-squared term depends on $v$, maximizing the likelihood
means minimizing chi-squared.

Most methods for radial-velocity estimation aren't explicitly maximizing
any likelihood.
Instead, they involve some kind of
cross-correlation of the data with a template.
This might be surprising, since (as we noted above) bound-saturating
estimators will be maximum-likelihood estimators.
However, under simple assumptions, cross-correlation optima are
also maximum-likelihood optima. The argument is as follows:
If we expand the quadratic form (chi-squared term) in the
likelihood in equation~(\ref{eq:lf}), it becomes
\begin{equation}
{}-\frac{1}{2}\,y\T\cdot C^{-1}\cdot y + y^T\cdot C^{-1}\cdot f(v) 
-\frac{1}{2}\,f(v)\T\cdot C^{-1}\cdot f(v) \quad.
\end{equation}
That is, the chi-squared term factors into a term that is the square
of the data $y$, a term that is the square of the model $f(v)$,
and a term that is a cross-correlation of data times model.
As long as the amplitude of $f(v)$ doesn't depend strongly on velocity $v$,
and the $C$ tensor is ``smooth'' in certain respects, the dependence of this
chi-squared term on velocity will be dominated by the
cross-correlation.
For this reason, we can use a variant of cross-correlation
optimization as a radial-velocity estimator.
